### Author: Ally Sprik
### Last-updated: 25-02-2024

Goal of this notebook is to impute missing values using Expectation Maximization (EM) and the expert defined network. The imputed data will be used to fit the pre-existing network. Define your postfix


```{r}
# Load libraries
library(bnlearn)
library(Rgraphviz)
library(gRain)
library(parallel)


postfix = "952"
```



```{r}
# Load the data set
df <- read.csv("0.2. Cleaned_data/Training_JAMA+Brno_model_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "NaN", "NULL", "", " ", "?"))
# uncomment the following line to only include the patients that are included in the original training cohort
#df <- df[df$Included_in_training_cohort == "yes",]

# Convert the data to factors, making them categorical
df <- as.data.frame(lapply(df, as.factor))

# Change the name of the p53 column to lowercase
names(df)[names(df) == 'P53'] <- 'p53'

# Define the Directed Acyclic Graph (DAG) structure
dag = paste("[MyometrialInvasion|PostoperativeGrade:Therapy][MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[Histology|FIGO:PreoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Recurrence_location|Recurrence]",
            "[Therapy|PostoperativeGrade]",
            "[Chemotherapy|Therapy][Radiotherapy|Therapy]",
            "[Survival1yr|Recurrence:Therapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Therapy:Survival5yr]",
            "[Survival5yr|Recurrence:Therapy]", sep="")


# convert the string to a network
BN <- model2network(dag)

# Show the DAG
graphviz.plot(BN)
```



Select the nodes in the data
```{r}
x <- as.array(bnlearn::nodes(BN))
df_select <- df[,x]
```


Impute the missing values using Expectation Maximization (EM)
```{r}
# Create cluster
cl <- makeCluster(detectCores() - 1)

# impute the missing values through Expectation maximization
BN <- bn.fit(BN, data = df_select, method = "bayes")

StructureEM <- structural.em(df_select, max.iter = 100, return.all = TRUE, maximize = 'tabu', start = BN)
```

unpack the data from EM
```{r}
# Unpack the data
dag <- StructureEM$dag
imputed <- StructureEM$imputed
Fitted <- StructureEM$fitted

# Add the extra columns to imputed
imputed <- cbind(imputed, df[,!(names(df) %in% names(imputed))])

```

```{r}
# Plot the DAG
graphviz.plot(dag)
```

Save the imputed data
```{r}
write.table(imputed, file = paste0("0. Source_files/0.3. Imputed_data/Training_JAMA_EM_imputed_",postfix,".csv"), sep = ",", row.names = FALSE)

```


Save the network
```{r}
write.net(Fitted, file = paste0("1. Ally_project/1.3. Model/Fitted_Networks/EM_fitted_DAG_",postfix,".net"))
```

