---
### Author: Ally Sprik
### Last-updated: 25-02-2024

Goal of this notebook is to generate networks in a variety of TCGA configurations, without platelets

description: Create 4 different Bayesian Networks for the TCGA data
            1. POLE to LNM and PostoperativeGrade
            2. POLE to PostoperativeGrade and Recurrence
            3. POLE to LNM and Recurrence
            4. POLE to LNM, PostoperativeGrade and Recurrence
---

```{r}
library(bnlearn)
library(Rgraphviz)

postfix = "952"

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph
# Define the model string for Dag_1
dag_1 = paste("[MyometrialInvasion|PostoperativeGrade]",#[MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")

# dagmodel: pre-defined directed acyclic graph (DAG) structure version 1
dagmodel = model2network(dag_1)

# Plot the Graphs
graphviz.plot(dagmodel, shape = "rectangle", highlight = list("nodes"))
```

# Import the data and and fit the model

```{r}
# Nodes in the model
nodesDag = names(dagmodel$nodes)

# Import the data
data = read.csv("0.1. Cleaned_data/Training+TCGA+JAMA_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))

# Select only the rows that have yes in Included_in_training_cohort
#data = data[data$Included_in_training_cohort == "yes",]
# Select for nodes in model
data = as.data.frame(lapply(data[, nodesDag], as.factor))

```

# Fit the model

```{r}
# Fit the model
fittedModel = bn.fit(dagmodel, data = data, method = 'bayes')
fittedNet = bn.net(fittedModel)
imputedData = impute(fittedModel, data, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

```
```{r}
# Load and insert CPTs
combined_MSI_POLE = read.csv("1. Ally_project/1.3. Model/CPT_files/MSI_POLE_combined.csv", header = TRUE, sep = ",")
combined_FIGO = read.csv("1. Ally_project/1.3. Model/CPT_files/FIGO_combined.csv", header = TRUE, sep = ",")

# Compute new MSI CPT
countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countMSI = countMSI + laplace_constant
cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

fittedModeli$MSI = as.table(cptMSIsmoothed)

# Compute new POLE CPT
countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countPOLE = countPOLE + laplace_constant
cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

names(dimnames(cptPOLEsmoothed)) = c("POLE", "LNM", "PostoperativeGrade")

fittedModeli$POLE = as.table(cptPOLEsmoothed)

# Compute new FIGO CPT
countFIGO = xtabs(~ FIGO + PostoperativeGrade, combined_FIGO)
laplace_constant = 1
smoothed_countFIGO = countFIGO + laplace_constant
cptFIGOsmoothed = prop.table(smoothed_countFIGO, 2)

names(dimnames(cptFIGOsmoothed)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGOsmoothed)

```
```{r}

# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

url = paste("1. Ally_project/1.3. Model/Fitted_Networks/R_WOP_AddTCGA_1_fitted_", postfix, ".net", sep="")
write.net(url, fittedModeli)

```

Everything below here doesn't currently work with the new way the CPTs are inserted, due to different CPTs count tables necessary. Thus it has been commented out.
Not a big deal to fix, you could load in the data and create local CPTs here instead of using the CSVs which is probably the best way. However, we found that for LNM the configurations didn't change anything, perhaps when looking at the survival side it can be interesting to see if the configurations change anything.
Currently, configuration 1, which is also the same configuration used in the other models that contain TCGA.


[//]: # ()
[//]: # (Now for the second arrangement of the nodes)

[//]: # ()
[//]: # (```{r})

[//]: # (# Define the model string for Dag_1)

[//]: # (dag_2 = paste&#40;"[MyometrialInvasion|PostoperativeGrade]",#[MRI_MI|MyometrialInvasion]",)

[//]: # (            "[Cytology|PostoperativeGrade]",)

[//]: # (            "[PreoperativeGrade]",)

[//]: # (            "[PostoperativeGrade|PreoperativeGrade]","[FIGO|PostoperativeGrade]",)

[//]: # (            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",)

[//]: # (            "[CA125|LNM][CTMRI|LNM:CA125]",)

[//]: # (            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",)

[//]: # (            "[p53|LNM:PostoperativeGrade:L1CAM]",)

[//]: # (            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",)

[//]: # (            "[L1CAM|LNM:PostoperativeGrade:PR]",)

[//]: # (            "[MSI|LNM:PostoperativeGrade]",)

[//]: # (            "[POLE|Recurrence:PostoperativeGrade]",)

[//]: # (            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",)

[//]: # (            "[Chemotherapy|PostoperativeGrade]",)

[//]: # (            "[Radiotherapy|MyometrialInvasion]",)

[//]: # (            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",)

[//]: # (            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",)

[//]: # (            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep=""&#41;)

[//]: # ()
[//]: # (# dagmodel: pre-defined directed acyclic graph &#40;DAG&#41; structure version 1)

[//]: # (dagmodel = model2network&#40;dag_2&#41;)

[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;dagmodel, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # ()
[//]: # (```)

[//]: # ()
[//]: # ()
[//]: # (Fit the model)

[//]: # ()
[//]: # (```{r})

[//]: # (# Fit the model)

[//]: # (fittedModel = bn.fit&#40;dagmodel, data = data, method = 'bayes'&#41;)

[//]: # (fittedNet = bn.net&#40;fittedModel&#41;)

[//]: # (imputedData = impute&#40;fittedModel, data, method = "bayes-lw", n=5000&#41;)

[//]: # (fittedModeli = bn.fit&#40;dagmodel, data = imputedData&#41;)

[//]: # (```)

[//]: # (```{r})

[//]: # (# Load and insert CPTs)

[//]: # (combined_MSI_POLE = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/MSI_POLE_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # (combined_FIGO = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/FIGO_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # ()
[//]: # (# Compute new MSI CPT)

[//]: # (countMSI = xtabs&#40;~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countMSI = countMSI + laplace_constant)

[//]: # (cptMSIsmoothed = prop.table&#40;smoothed_countMSI, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptMSIsmoothed&#41;&#41; = c&#40;"MSI", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$MSI = as.table&#40;cptMSIsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new POLE CPT)

[//]: # (countPOLE = xtabs&#40;~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countPOLE = countPOLE + laplace_constant)

[//]: # (cptPOLEsmoothed = prop.table&#40;smoothed_countPOLE, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptPOLEsmoothed&#41;&#41; = c&#40;"POLE", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$POLE = as.table&#40;cptPOLEsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new FIGO CPT)

[//]: # (countFIGO = xtabs&#40;~ FIGO + PostoperativeGrade, combined_FIGO&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countFIGO = countFIGO + laplace_constant)

[//]: # (cptFIGOsmoothed = prop.table&#40;smoothed_countFIGO, 2&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptFIGOsmoothed&#41;&#41; = c&#40;"FIGO", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$FIGO = as.table&#40;cptFIGOsmoothed&#41;)

[//]: # ()
[//]: # (```)

[//]: # (```{r})

[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;fittedNet, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # (url = paste&#40;"1. Ally_project/1.3. Model/Fitted_Networks/R_WOP_AddTCGA_2_fitted_", postfix, ".net", sep=""&#41;)

[//]: # (write.net&#40;url, fittedModeli&#41;)

[//]: # ()
[//]: # (```)

[//]: # ()
[//]: # ()
[//]: # (Now for the third arrangement of the nodes)

[//]: # ()
[//]: # (```{r})

[//]: # (# Define the model string for Dag_1)

[//]: # (dag_3 = paste&#40;"[MyometrialInvasion|PostoperativeGrade]",#[MRI_MI|MyometrialInvasion]",)

[//]: # (            "[Cytology|PostoperativeGrade]",)

[//]: # (            "[PreoperativeGrade]",)

[//]: # (            "[PostoperativeGrade|PreoperativeGrade]", "[FIGO|PostoperativeGrade]",)

[//]: # (            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",)

[//]: # (            "[CA125|LNM][CTMRI|LNM:CA125]",)

[//]: # (            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",)

[//]: # (            "[p53|LNM:PostoperativeGrade:L1CAM]",)

[//]: # (            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",)

[//]: # (            "[L1CAM|LNM:PostoperativeGrade:PR]",)

[//]: # (            "[MSI|LNM:PostoperativeGrade]",)

[//]: # (            "[POLE|Recurrence:LNM]",)

[//]: # (            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",)

[//]: # (            "[Chemotherapy|PostoperativeGrade]",)

[//]: # (            "[Radiotherapy|MyometrialInvasion]",)

[//]: # (            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",)

[//]: # (            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",)

[//]: # (            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep=""&#41;)

[//]: # ()
[//]: # (# dagmodel: pre-defined directed acyclic graph &#40;DAG&#41; structure version 1)

[//]: # (dagmodel = model2network&#40;dag_3&#41;)

[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;dagmodel, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # ()
[//]: # (```)

[//]: # ()
[//]: # ()
[//]: # (Fit the model)

[//]: # ()
[//]: # (```{r})

[//]: # (# Fit the model)

[//]: # (fittedModel = bn.fit&#40;dagmodel, data = data, method = 'bayes'&#41;)

[//]: # (fittedNet = bn.net&#40;fittedModel&#41;)

[//]: # (imputedData = impute&#40;fittedModel, data, method = "bayes-lw", n=5000&#41;)

[//]: # (fittedModeli = bn.fit&#40;dagmodel, data = imputedData&#41;)

[//]: # ()
[//]: # (```)

[//]: # (```{r})

[//]: # (# Load and insert CPTs)

[//]: # (combined_MSI_POLE = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/MSI_POLE_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # (combined_FIGO = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/FIGO_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # ()
[//]: # (# Compute new MSI CPT)

[//]: # (countMSI = xtabs&#40;~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countMSI = countMSI + laplace_constant)

[//]: # (cptMSIsmoothed = prop.table&#40;smoothed_countMSI, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptMSIsmoothed&#41;&#41; = c&#40;"MSI", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$MSI = as.table&#40;cptMSIsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new POLE CPT)

[//]: # (countPOLE = xtabs&#40;~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countPOLE = countPOLE + laplace_constant)

[//]: # (cptPOLEsmoothed = prop.table&#40;smoothed_countPOLE, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptPOLEsmoothed&#41;&#41; = c&#40;"POLE", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$POLE = as.table&#40;cptPOLEsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new FIGO CPT)

[//]: # (countFIGO = xtabs&#40;~ FIGO + PostoperativeGrade, combined_FIGO&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countFIGO = countFIGO + laplace_constant)

[//]: # (cptFIGOsmoothed = prop.table&#40;smoothed_countFIGO,2&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptFIGOsmoothed&#41;&#41; = c&#40;"FIGO", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$FIGO = as.table&#40;cptFIGOsmoothed&#41;)

[//]: # ()
[//]: # (```)

[//]: # (```{r})

[//]: # ()
[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;fittedNet, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # (url = paste&#40;"1. Ally_project/1.3. Model/Fitted_Networks/R_WOP_AddTCGA_3_fitted_", postfix, ".net", sep=""&#41;)

[//]: # (write.net&#40;url, fittedModeli&#41;)

[//]: # ()
[//]: # (```)

[//]: # ()
[//]: # ()
[//]: # ()
[//]: # (Now for the fourth arrangement of the nodes)

[//]: # ()
[//]: # (```{r})

[//]: # (# Define the model string for Dag_1)

[//]: # (dag_4 = paste&#40;"[MyometrialInvasion|PostoperativeGrade]",#[MRI_MI|MyometrialInvasion]",)

[//]: # (            "[Cytology|PostoperativeGrade]",)

[//]: # (            "[PreoperativeGrade]",)

[//]: # (            "[PostoperativeGrade|PreoperativeGrade]", "[FIGO|PostoperativeGrade]",)

[//]: # (            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",)

[//]: # (            "[CA125|LNM][CTMRI|LNM:CA125]",)

[//]: # (            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",)

[//]: # (            "[p53|LNM:PostoperativeGrade:L1CAM]",)

[//]: # (            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",)

[//]: # (            "[L1CAM|LNM:PostoperativeGrade:PR]",)

[//]: # (            "[MSI|LNM:PostoperativeGrade]",)

[//]: # (            "[POLE|Recurrence:LNM:PostoperativeGrade]",)

[//]: # (            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",)

[//]: # (            "[Chemotherapy|PostoperativeGrade]",)

[//]: # (            "[Radiotherapy|MyometrialInvasion]",)

[//]: # (            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",)

[//]: # (            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",)

[//]: # (            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep=""&#41;)

[//]: # ()
[//]: # (# dagmodel: pre-defined directed acyclic graph &#40;DAG&#41; structure version 1)

[//]: # (dagmodel = model2network&#40;dag_4&#41;)

[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;dagmodel, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # ()
[//]: # (```)

[//]: # ()
[//]: # ()
[//]: # (Fit the model)

[//]: # ()
[//]: # (```{r})

[//]: # (# Fit the model)

[//]: # (fittedModel = bn.fit&#40;dagmodel, data = data, method = 'bayes'&#41;)

[//]: # (fittedNet = bn.net&#40;fittedModel&#41;)

[//]: # (imputedData = impute&#40;fittedModel, data, method = "bayes-lw", n=5000&#41;)

[//]: # (fittedModeli = bn.fit&#40;dagmodel, data = imputedData&#41;)

[//]: # ()
[//]: # (```)

[//]: # (```{r})

[//]: # (# Load and insert CPTs)

[//]: # (combined_MSI_POLE = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/MSI_POLE_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # (combined_FIGO = read.csv&#40;"1. Ally_project/1.3. Model/CPT_files/FIGO_combined.csv", header = TRUE, sep = ","&#41;)

[//]: # ()
[//]: # (# Compute new MSI CPT)

[//]: # (countMSI = xtabs&#40;~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countMSI = countMSI + laplace_constant)

[//]: # (cptMSIsmoothed = prop.table&#40;smoothed_countMSI, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptMSIsmoothed&#41;&#41; = c&#40;"MSI", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$MSI = as.table&#40;cptMSIsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new POLE CPT)

[//]: # (countPOLE = xtabs&#40;~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countPOLE = countPOLE + laplace_constant)

[//]: # (cptPOLEsmoothed = prop.table&#40;smoothed_countPOLE, c&#40;3, 2&#41;&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptPOLEsmoothed&#41;&#41; = c&#40;"POLE", "LNM", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$POLE = as.table&#40;cptPOLEsmoothed&#41;)

[//]: # ()
[//]: # (# Compute new FIGO CPT)

[//]: # (countFIGO = xtabs&#40;~ FIGO + PostoperativeGrade, combined_FIGO&#41;)

[//]: # (laplace_constant = 1)

[//]: # (smoothed_countFIGO = countFIGO + laplace_constant)

[//]: # (cptFIGOsmoothed = prop.table&#40;smoothed_countFIGO, 2&#41;)

[//]: # ()
[//]: # (names&#40;dimnames&#40;cptFIGOsmoothed&#41;&#41; = c&#40;"FIGO", "PostoperativeGrade"&#41;)

[//]: # ()
[//]: # (fittedModeli$FIGO = as.table&#40;cptFIGOsmoothed&#41;)

[//]: # ()
[//]: # ()
[//]: # (```)

[//]: # (```{r})

[//]: # ()
[//]: # (# Plot the Graphs)

[//]: # (graphviz.plot&#40;fittedNet, shape = "rectangle", highlight = list&#40;"nodes"&#41;&#41;)

[//]: # ()
[//]: # (url = paste&#40;"1. Ally_project/1.3. Model/Fitted_Networks/R_WOP_AddTCGA_4_fitted_", postfix, ".net", sep=""&#41;)

[//]: # (write.net&#40;url, fittedModeli&#41;)

```