### Author: Ally Sprik
### Last-updated: 25-02-2024

Goal of this notebook is to generate networks with platelets and with and without MRI, MSI and POLE



```{r}
library(bnlearn)
library(dplyr)
library(gRain)

# Define postfix for filenames
postfix = "952+Brno"

# Define Directed Acyclic Graph (DAG) structure
# It is a string that defines the structure of the Bayesian network according to the following rules:
# [node1|node2:node3] means that node1 is dependent on node2 and node3
# A node must be defined before it can be used as a parent node
# each bracket pair defines a node and its parents
dag = paste("[MyometrialInvasion|PostoperativeGrade]",#MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            #"[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            #"[MSI|LNM:PostoperativeGrade]",
            #"[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# Transform the model string to Directed Acyclic Graph
dagmodel = model2network(dag)
# Nodes in the model
nodesDag = names(dagmodel$nodes)

# Import the data
#data = read.csv("0. Source_files/0.2. Cleaned_data/Training+TCGA+JAMA_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA",
# "N/A", "N/A ", " ", ""))

data = read.csv("0. Source_files/0.2. Cleaned_data/Training_JAMA+Brno_model_complete_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))

# Select only the rows that have yes in Included_in_training_cohort
#data = data[data$Included_in_training_cohort == "yes",]
# Select for nodes in model
tdata = as.data.frame(lapply(data[, nodesDag], as.factor))

```


# Fit the model
- Fit the model on the training data
- Impute the missing data
- Fit the model on the imputed data
```{r}
fittedModel = bn.fit(dagmodel, data = tdata, method = 'bayes')
imputedData = impute(fittedModel, tdata, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

fittedNet = bn.net(fittedModeli)
# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

url = paste("1. Ally_project/1.3. Model/Fitted_Networks/R_WP_",postfix,".net", sep = "")
write.net(url, fittedModeli)
```


Next model is the same as the previous model but with MRI, MSI and POLE
```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade][MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# dagmodel: pre-defined directed acyclic graph (DAG) structure
dagmodel = model2network(dag)

# Plot the Graphs
graphviz.plot(dagmodel, shape = "rectangle", highlight = list("nodes"))

```

# Import the data and and fit the model

```{r}
# Nodes in the model
nodesDag = names(dagmodel$nodes)

# Select for nodes in model
tdata = as.data.frame(lapply(data[, nodesDag], as.factor))

```

# Fit the model

```{r}
# Fit the model
fittedModel = bn.fit(dagmodel, data = tdata, method = 'bayes')
imputedData = impute(fittedModel, tdata, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
fittedNet = bn.net(fittedModel)
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))
```

Calculate and insert the local CPTs

Pseudocode:
1. Create a subset of the data with the nodes of interest
2. Remove the rows with missing values
3. Repeat for each node of interest
```{r}
combined_MRI_MI = data.frame(MRI_MI = data$MRI_MI, MyometrialInvasion = imputedData$MyometrialInvasion)
combined_MRI_MI <- combined_MRI_MI[complete.cases(combined_MRI_MI), ]

combined_MSI_POLE = data.frame(MSI = data$MSI, POLE=data$POLE , LNM = imputedData$LNM, PostoperativeGrade = imputedData$PostoperativeGrade)
combined_MSI_POLE <- combined_MSI_POLE[complete.cases(combined_MSI_POLE), ]

combined_FIGO = data.frame(FIGO = data$FIGO, PostoperativeGrade = imputedData$PostoperativeGrade)
combined_FIGO <- combined_FIGO[complete.cases(combined_FIGO), ]

```

Save the data for the calculation of the local CPTs
```{r}

# Save the data for the calculation of the local CPTs
write.csv(combined_MRI_MI, "1. Ally_project/1.3. Model/CPT_files/MRI_MI_combined.csv",  row.names = FALSE)
write.csv(combined_MSI_POLE, "1. Ally_project/1.3. Model/CPT_files/MSI_POLE_combined.csv",  row.names = FALSE)
write.csv(combined_FIGO, "1. Ally_project/1.3. Model/CPT_files/FIGO_combined.csv",  row.names = FALSE)

```

Calculate and insert the local CPTs
Pseudocode:
1. Create a subset of the data with the nodes of interest
2. Calculate the CPT
3. Ensure the names of the dimensions are correct
4. Insert the CPT in the model
5. Repeat for each node of interest

```{r}
# FIGO
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MRI CPT
countMRI = table(combined_MRI_MI$MRI_MI, combined_MRI_MI$MyometrialInvasion)
cptMRI = prop.table(countMRI, 2)

names(dimnames(cptMRI)) = c("MRI_MI", "MyometrialInvasion")

fittedModeli$MRI_MI = as.table(cptMRI)

# Compute new MSI CPT
countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countMSI = countMSI + laplace_constant
cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

# names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

fittedModeli$MSI = as.table(cptMSIsmoothed)

# Compute new POLE CPT
countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countPOLE = countPOLE + laplace_constant
cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

# names(dimnames(cptPOLEsmoothed)) = c("POLE", "PostoperativeGrade", "LNM")

fittedModeli$POLE = as.table(cptPOLEsmoothed)

```

Save the model
```{r}

url = paste("1. Ally_project/1.3. Model/Fitted_Networks/R_WP_all_",postfix,".net", sep = "")
write.net(url, fittedModeli)

url = paste("0.3. Imputed_data/R_WP_all_model_imputed_",postfix,".csv", sep = "")
write.csv(imputedData, url,  row.names = FALSE)
```

Next model is the same as the previous model but with MRI_MI without MSI and POLE

```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade][MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            #"[MSI|LNM:PostoperativeGrade]",
            #"[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# Confert the modelstring to a Directed Acyclic Graph
dagmodel = model2network(dag)
# Select for nodes in model
tempdata = data[, names(dagmodel$nodes)]
# Tempdata as factor
tempdata = as.data.frame(lapply(tempdata, as.factor))
```


```{r}
# Nodes in the model
# Fit the model on the training data
fittedModel = bn.fit(dagmodel, data = tempdata, method = 'bayes')
# Impute the missing data
imputedData = impute(fittedModel, tempdata, method = "bayes-lw", n=5000)
# Fit the model on the imputed data
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
fittedNet = bn.net(fittedModel)
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

```


Pseudocode:
1. Create a subset of the data with the nodes of interest
2. calculate the CPT
3. Ensure the names of the dimensions are correct
4. Insert the CPT in the model
```{r}
# Calculate and insert the local CPTs
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MRI CPT
countMRI = table(combined_MRI_MI$MRI_MI, combined_MRI_MI$MyometrialInvasion)
cptMRI = prop.table(countMRI, 2)

names(dimnames(cptMRI)) = c("MRI_MI", "MyometrialInvasion")

fittedModeli$MRI_MI = as.table(cptMRI)

```

# save the model
```{r}
url = paste("1. Ally_project/1.3. Model/Fitted_Networks/R_WP_MRI_",postfix,".net", sep = "")
write.net(url, fittedModeli)
```


Next model is the same as the previous model but with MSI and POLE

```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade]",# [MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# Confert the modelstring to a Directed Acyclic Graph
dagmodel = model2network(dag)
# Select for nodes in model
tempdata = data[, names(dagmodel$nodes)]
# Tempdata as factor
tempdata = as.data.frame(lapply(tempdata, as.factor))
```


```{r}
# Nodes in the model
# Fit the model
fittedModel = bn.fit(dagmodel, data = tempdata, method = 'bayes')
# Impute the missing data
imputedData = impute(fittedModel, tempdata, method = "bayes-lw", n=5000)
# Fit the model on the imputed data
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
fittedNet = bn.net(fittedModel)
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

# Calculate and insert the local CPTs as described above
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MSI CPT
countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countMSI = countMSI + laplace_constant
cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

#names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

fittedModeli$MSI = as.table(cptMSIsmoothed)

# Compute new POLE CPT
countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countPOLE = countPOLE + laplace_constant
cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

#names(dimnames(cptPOLEsmoothed)) = c("POLE", "PostoperativeGrade", "LNM")

fittedModeli$POLE = as.table(cptPOLEsmoothed)

```

# save the model
```{r}

url = paste("1. Ally_project/1.3. Model/Fitted_Networks/R_WP_TCGA_",postfix,".net", sep = "")
write.net(url, fittedModeli)
```



