### Author: Ally Sprik
### Last-updated: 25-02-2024

Goal of this notebook is to generate networks with platelets and with and without MRI, MSI and POLE



```{r}
library(bnlearn)
library(dplyr)
library(gRain)

dag = paste("[MyometrialInvasion|PostoperativeGrade]",#MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            #"[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            #"[MSI|LNM:PostoperativeGrade]",
            #"[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# dagmodel: pre-defined directed acyclic graph (DAG) structure
dagmodel = model2network(dag)
# Nodes in the model
nodesDag = names(dagmodel$nodes)

# Import the data
data = read.csv("0.1. Cleaned_data/Training_JAMA+Brno_model_complete_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))
# Select only the rows that have yes in Included_in_training_cohort
#data = data[data$Included_in_training_cohort == "yes",]
# Select for nodes in model
data = as.data.frame(lapply(data[, nodesDag], as.factor))

```


# Fit the model
```{r}
fittedModel = bn.fit(dagmodel, data = data, method = 'bayes')
fittedNet = bn.net(fittedModel)
imputedData = impute(fittedModel, data, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))
write.net("3. Model/Fitted_Networks/R_WP_original_952+Brno.net", fittedModeli)
```


Next model is the same as the previous model but with MRI, MSI and POLE
```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade][MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# dagmodel: pre-defined directed acyclic graph (DAG) structure
dagmodel = model2network(dag)

# Plot the Graphs
graphviz.plot(dagmodel, shape = "rectangle", highlight = list("nodes"))

```

# Import the data and and fit the model

```{r}
# Nodes in the model
nodesDag = names(dagmodel$nodes)

# Import the data
data = read.csv("0.1. Cleaned_data/Training+TCGA+JAMA_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))
# Select only the rows that have yes in Included_in_training_cohort
data = data[data$Included_in_training_cohort == "yes",]
# Select for nodes in model
data = as.data.frame(lapply(data[, nodesDag], as.factor))

```

# Fit the model

```{r}
# Fit the model
fittedModel = bn.fit(dagmodel, data = data, method = 'bayes')
fittedNet = bn.net(fittedModel)
imputedData = impute(fittedModel, data, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))
```

Calculate and insert the local CPTs
```{r}
data = read.csv("0.1. Cleaned_data/Training+TCGA+JAMA_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))
data = data[data$Included_in_training_cohort == "yes",]
# Combine the data for the calculation of the local CPTs
combined_MRI_MI = data.frame(MRI_MI = data$MRI_MI, MyometrialInvasion = imputedData$MyometrialInvasion)
combined_MRI_MI <- combined_MRI_MI[complete.cases(combined_MRI_MI), ]

combined_MSI_POLE = data.frame(MSI = data$MSI, POLE=data$POLE , LNM = imputedData$LNM, PostoperativeGrade = imputedData$PostoperativeGrade)
combined_MSI_POLE <- combined_MSI_POLE[complete.cases(combined_MSI_POLE), ]

combined_FIGO = data.frame(FIGO = data$FIGO, PostoperativeGrade = imputedData$PostoperativeGrade)
combined_FIGO <- combined_FIGO[complete.cases(combined_FIGO), ]

```

Save the data for the calculation of the local CPTs
```{r}

# Save the data for the calculation of the local CPTs
write.csv(combined_MRI_MI, "3. Model/CPT_files/MRI_MI_combined.csv",  row.names = FALSE)
write.csv(combined_MSI_POLE, "3. Model/CPT_files/MSI_POLE_combined.csv",  row.names = FALSE)
write.csv(combined_FIGO, "3. Model/CPT_files/FIGO_combined.csv",  row.names = FALSE)

```

Calculate and insert the local CPTs
```{r}
# Calculate and insert the local CPTs
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MRI CPT
countMRI = table(combined_MRI_MI$MRI_MI, combined_MRI_MI$MyometrialInvasion)
cptMRI = prop.table(countMRI, 2)

names(dimnames(cptMRI)) = c("MRI_MI", "MyometrialInvasion")

fittedModeli$MRI_MI = as.table(cptMRI)

# Compute new MSI CPT
countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countMSI = countMSI + laplace_constant
cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

# names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

fittedModeli$MSI = as.table(cptMSIsmoothed)

# Compute new POLE CPT
countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countPOLE = countPOLE + laplace_constant
cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

# names(dimnames(cptPOLEsmoothed)) = c("POLE", "PostoperativeGrade", "LNM")

fittedModeli$POLE = as.table(cptPOLEsmoothed)

```

Save the model
```{r}
write.net("3. Model/Fitted_Networks/R_WP_all_952+Brno.net", fittedModeli)

write.csv(imputedData, "0.2. Imputed_data/Training_JAMA+Brno_imputed_952+Brno.csv",  row.names = FALSE)
```

Next model is the same as the previous model but with MRI_MI without MSI and POLE

```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade][MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            #"[MSI|LNM:PostoperativeGrade]",
            #"[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# dagmodel: pre-defined directed acyclic graph (DAG) structure
dagmodel = model2network(dag)
tempdata = data[, names(dagmodel$nodes)]
# Tempdata as factor
tempdata = as.data.frame(lapply(tempdata, as.factor))
```


```{r}
# Nodes in the model
# Fit the model
fittedModel = bn.fit(dagmodel, data = tempdata, method = 'bayes')
fittedNet = bn.net(fittedModel)
imputedData = impute(fittedModel, tempdata, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

# Calculate and insert the local CPTs
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MRI CPT
countMRI = table(combined_MRI_MI$MRI_MI, combined_MRI_MI$MyometrialInvasion)
cptMRI = prop.table(countMRI, 2)

names(dimnames(cptMRI)) = c("MRI_MI", "MyometrialInvasion")

fittedModeli$MRI_MI = as.table(cptMRI)

```

# save the model
```{r}
write.net("3. Model/Fitted_Networks/R_WP_MRI_952+Brno.net", fittedModeli)
```


Next model is the same as the previous model but with MSI and POLE

```{r}
library(bnlearn)
library(Rgraphviz)

# The necessary subsets are created in the data-cleaning.R
# Transform the modelstring (bayesian network) to Directed Acyclic Graph

dag = paste("[MyometrialInvasion|PostoperativeGrade]",# [MRI_MI|MyometrialInvasion]",
            "[Cytology|PostoperativeGrade]",
            "[PreoperativeGrade]",
            "[PostoperativeGrade|PreoperativeGrade]",
            "[FIGO|PostoperativeGrade]",
            "[LNM|LVSI:MyometrialInvasion:PostoperativeGrade:Chemotherapy]",
            "[CA125|LNM][CTMRI|LNM:CA125]",
            "[LVSI|PostoperativeGrade:Cytology:Chemotherapy]",
            "[p53|LNM:PostoperativeGrade:L1CAM]",
            "[ER|LNM:PostoperativeGrade][PR|LNM:PostoperativeGrade:ER]",
            "[L1CAM|LNM:PostoperativeGrade:PR]",
            "[Platelets|LNM]",
            "[MSI|LNM:PostoperativeGrade]",
            "[POLE|LNM:PostoperativeGrade]",
            "[Recurrence|CA125:LNM:LVSI:PostoperativeGrade]",
            "[Chemotherapy|PostoperativeGrade]",
            "[Radiotherapy|MyometrialInvasion]",
            "[Survival1yr|Recurrence:Chemotherapy:Radiotherapy:Survival3yr:Survival5yr]",
            "[Survival3yr|Recurrence:Chemotherapy:Radiotherapy:Survival5yr]",
            "[Survival5yr|Recurrence:Chemotherapy:Radiotherapy]", sep="")


# dagmodel: pre-defined directed acyclic graph (DAG) structure
dagmodel = model2network(dag)
tempdata = data[, names(dagmodel$nodes)]
tempdata = as.data.frame(lapply(tempdata, as.factor))
```


```{r}
# Nodes in the model
# Fit the model
fittedModel = bn.fit(dagmodel, data = tempdata, method = 'bayes')
fittedNet = bn.net(fittedModel)
imputedData = impute(fittedModel, tempdata, method = "bayes-lw", n=5000)
fittedModeli = bn.fit(dagmodel, data = imputedData)

# Plot the Graphs
graphviz.plot(fittedNet, shape = "rectangle", highlight = list("nodes"))

# Calculate and insert the local CPTs
countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
cptFIGO = prop.table(countFIGO, 2)

names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

fittedModeli$FIGO = as.table(cptFIGO)

# Compute new MSI CPT
countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countMSI = countMSI + laplace_constant
cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

# names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

fittedModeli$MSI = as.table(cptMSIsmoothed)

# Compute new POLE CPT
countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
laplace_constant = 1
smoothed_countPOLE = countPOLE + laplace_constant
cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

# names(dimnames(cptPOLEsmoothed)) = c("POLE", "PostoperativeGrade", "LNM")

fittedModeli$POLE = as.table(cptPOLEsmoothed)

```

# save the model
```{r}
write.net("3. Model/Fitted_Networks/R_WP_TCGA_952+Brno.net", fittedModeli)
```



