Author: Ally Sprik
Date: 25-01-2024

Summary:
In this file the Brno Dataset will be loaded and combined with the training set in KFolds, the model will then be trained and evaluated on each of the folds. As well as comparing the results to the results of the model trained on the training set only on each of the Folds. The results will be saved in a csv file for each of the folds, to be used in python for analysis.


### Load Libraries and Data

```{r}
library(tidyverse)
library(caret)
library(bnlearn)
library(dplyr)
library(Rgraphviz)
library(gRain)

# Load the data
df_or <- read.csv("0.1. Cleaned_data/Training+TCGA+JAMA_cleaned.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))
df_brno <- read.csv("0.1. Cleaned_data/Cleaned_Brno_model_complete.csv", header = TRUE, sep = ",", na.strings = c("NA", "N/A", "N/A ", " ", ""))
# change name FIGO surgical to FIGO
colnames(df_brno)[colnames(df_brno) == "FIGO_surgical"] <- "FIGO"

# Create NA columns for the columns that are not in the Brno dataset
df_brno["CTMRI"] <- NA

# Get the columns that are in both datasets
common_cols <- intersect(colnames(df_or), colnames(df_brno))


df_or <- df_or[,common_cols]
df_brno <- df_brno[,common_cols]

# Load networks
dag_Or <- model2network(modelstring(read.net("0.3. Original_Casper_files/Results/Casper_fitted_952+Brno.net")))
dag_NoPlat_NoRest <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WOP_original_952+Brno.net")))
dag_NoPlat_MRI <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WOP_AddMRIMI_fitted_952+Brno.net")))
dag_NoPlat_TCGA <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WOP_AddTCGA_1_fitted_952+Brno.net")))
dag_NoPlat_TCGA_MRI <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WOP_AddMRIMITCGA_fitted_952+Brno.net")))
dag_Plat_NoRest <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WP_original_952+Brno.net")))
dag_Plat_TCGA_MRI <- model2network(modelstring(read.net("3. Model/Fitted_Networks/R_WP_all_952+Brno.net")))

```

```{r}
folds <- createFolds(df_brno$LNM, k =5, list=TRUE, returnTrain = TRUE)


```

```{r}

testsets <- list()
for (i in 1:6){
  test <- df_brno[-folds[[i]],]
  testsets[[i]] <- test
}

```

```{r}

getNetworks <- function(bNet, loc = 0, FIGO = 1, TCGA = 0, MRI = 0) {
  networks <- list()
  if (loc == 1){
    # Change the name of the column Recurrence to Recurrence_bi and Recurrence_location to Recurrence
    colnames(df_brno)[colnames(df_brno) == "Recurrence"] <- "Recurrence_bi"
    colnames(df_brno)[colnames(df_brno) == "Recurrence_location"] <- "Recurrence"
    # Change the name of the column Recurrence to Recurrence_bi and Recurrence_location to Recurrence
    colnames(df_or)[colnames(df_or) == "Recurrence"] <- "Recurrence_bi"
    colnames(df_or)[colnames(df_or) == "Recurrence_location"] <- "Recurrence"
  }

  for (i in 1:6){
    network <- bNet
    nodenames <- names(network$nodes)
    # Get the fold
    brno_fold <- df_brno[folds[[i]],]
    # concat brno fold with or training set
    data <- rbind(df_or, brno_fold)


    # Train and get the networks
    data <- as.data.frame(lapply(data[,nodenames], as.factor))

    fittedModel <- bnlearn::bn.fit(network, data = data, method = 'bayes', iss=2)
    imputedData <- bnlearn::impute(fittedModel, data, method = "bayes-lw", n=5000)
    fittedModeli <- bnlearn::bn.fit(network, data = imputedData, method = 'bayes', iss=2)


    if (FIGO == 1){
      combined_FIGO <- data.frame(FIGO = data$FIGO, PostoperativeGrade = imputedData$PostoperativeGrade)
      combined_FIGO <- combined_FIGO[complete.cases(combined_FIGO), ]

      # Calculate and insert the local CPTs
      countFIGO = table(combined_FIGO$FIGO, combined_FIGO$PostoperativeGrade)
      cptFIGO = prop.table(countFIGO, 2)

      names(dimnames(cptFIGO)) = c("FIGO", "PostoperativeGrade")

      fittedModeli$FIGO = as.table(cptFIGO)
    }

    if (TCGA == 1){
      combined_MSI_POLE <- data.frame(MSI = data$MSI, POLE=data$POLE , LNM = imputedData$LNM, PostoperativeGrade = imputedData$PostoperativeGrade)
      combined_MSI_POLE <- combined_MSI_POLE[complete.cases(combined_MSI_POLE), ]
      # Compute new MSI CPT
      countMSI = xtabs(~ MSI + LNM + PostoperativeGrade, combined_MSI_POLE)
      laplace_constant = 1
      smoothed_countMSI = countMSI + laplace_constant
      cptMSIsmoothed = prop.table(smoothed_countMSI, c(3, 2))

      # names(dimnames(cptMSIsmoothed)) = c("MSI", "LNM", "PostoperativeGrade")

      fittedModeli$MSI = as.table(cptMSIsmoothed)

      # Compute new POLE CPT
      countPOLE = xtabs(~ POLE + LNM + PostoperativeGrade, combined_MSI_POLE)
      laplace_constant = 1
      smoothed_countPOLE = countPOLE + laplace_constant
      cptPOLEsmoothed = prop.table(smoothed_countPOLE, c(3, 2))

      # names(dimnames(cptPOLEsmoothed)) = c("POLE", "PostoperativeGrade", "LNM")

      fittedModeli$POLE = as.table(cptPOLEsmoothed)
    }

    if (MRI == 1){
      combined_MRI_MI <- data.frame(MRI_MI = data$MRI_MI, MyometrialInvasion = imputedData$MyometrialInvasion)
      combined_MRI_MI <- combined_MRI_MI[complete.cases(combined_MRI_MI), ]

      # Compute new MRI CPT
      countMRI = table(combined_MRI_MI$MRI_MI, combined_MRI_MI$MyometrialInvasion)
      cptMRI = prop.table(countMRI, 2)

      names(dimnames(cptMRI)) = c("MRI_MI", "MyometrialInvasion")

      fittedModeli$MRI_MI = as.table(cptMRI)
    }

    networks[[i]] <- fittedModeli
  }

  # Change the names back if changed
    if (loc == 1){
        # Change the name of the column Recurrence to Recurrence_bi and Recurrence_location to Recurrence
        colnames(df_brno)[colnames(df_brno) == "Recurrence_bi"] <- "Recurrence"
        colnames(df_brno)[colnames(df_brno) == "Recurrence"] <- "Recurrence_location"
        # Change the name of the column Recurrence to Recurrence_bi and Recurrence_location to Recurrence
        colnames(df_or)[colnames(df_or) == "Recurrence_bi"] <- "Recurrence"
        colnames(df_or)[colnames(df_or) == "Recurrence"] <- "Recurrence_location"
    }

  return (networks)
}

```

```{r}
# Get the networks
print("Started 1")
networks_Or <- getNetworks(dag_Or, loc = 1, FIGO = 0)
print("Started 2")
networks_NoPlat_NoRest <- getNetworks(dag_NoPlat_NoRest)
print("Started 3")
networks_NoPlat_MRI <- getNetworks(dag_NoPlat_MRI, MRI = 1)
print("Started 4")
networks_NoPlat_TCGA <- getNetworks(dag_NoPlat_TCGA, TCGA = 1)
print("Started 5")
networks_NoPlat_TCGA_MRI <- getNetworks(dag_NoPlat_TCGA_MRI, TCGA = 1, MRI = 1)
print("Started 6")
networks_Plat_NoRest <- getNetworks(dag_Plat_NoRest, FIGO = 0)
print("Started 7")
networks_Plat_TCGA_MRI <- getNetworks(dag_Plat_TCGA_MRI, TCGA = 1, MRI = 1)

```


### Validation of the different sets

```{r}

evidence_columns <- c("PreoperativeGrade", "MRI_MI", "CA125", "Platelets", "L1CAM", "ER", "PR", "p53", "POLE", "MSI","CTMRI", "Histology", "Cytology")

getResults <- function(networks){
  results <- list()
  for (i in 1:6){
    # Get the fold
    test <- testsets[[i]]
    # Get the network
    network <- as.grain(networks[[i]])
    # Get the nodes
    nodenames <- names(network$cptlist)
    # Common set of nodenames and evidence columns
    common <- intersect(nodenames, evidence_columns)
    # Get the test data
    test <- as.data.frame(lapply(test[,common], as.factor))
    # Get the predictions
    predictions <- list()


    for (j in 1:nrow(test)){
      evidence <- as.list(test[j,])
      evidence <- as.data.frame(evidence[!is.na(evidence)])

      prediction <- predict(network, response =c("LNM","Survival5yr"), newdata = evidence, type="distribution") # Select the yes value from the nodes to make life easier

      predictions$LNM[[j]] <- prediction$pred$LNM[[2]]
      predictions$Surv[[j]] <- prediction$pred$Survival5yr[[2]]
    }
    results[[i]] <- predictions
    # Get the results
  }
  return (results)
}

```

```{r}
# Get the results
print("Started 1")
results_Or <- getResults(networks_Or)
print("Started 2")
results_NoPlat_NoRest <- getResults(networks_NoPlat_NoRest)
print("Started 3")
results_NoPlat_MRI <- getResults(networks_NoPlat_MRI)
print("Started 4")
results_NoPlat_TCGA <- getResults(networks_NoPlat_TCGA)
print("Started 5")
results_NoPlat_TCGA_MRI <- getResults(networks_NoPlat_TCGA_MRI)
print("Started 6")
results_Plat_NoRest <- getResults(networks_Plat_NoRest)
print("Started 7")
results_Plat_TCGA_MRI <- getResults(networks_Plat_TCGA_MRI)

```

```{r}
# Unpack the results
for (i in 1:6){
  # Save each of the results and the test set in a csv file, where the name is the result name + the fold number
  write.csv(results_Or[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_Or_LNM_", i, ".csv"))
  write.csv(results_Or[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_Or_Surv_", i, ".csv"))

  write.csv(results_NoPlat_NoRest[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_NoRest_LNM_", i, ".csv"))
  write.csv(results_NoPlat_NoRest[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_NoRest_Surv_", i, ".csv"))

  write.csv(results_NoPlat_MRI[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_MRI_LNM_", i, ".csv"))
  write.csv(results_NoPlat_MRI[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_MRI_Surv_", i, ".csv"))

  write.csv(results_NoPlat_TCGA[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_TCGA_LNM_", i, ".csv"))
  write.csv(results_NoPlat_TCGA[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_TCGA_Surv_", i, ".csv"))

  write.csv(results_NoPlat_TCGA_MRI[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_TCGA_MRI_LNM_", i, ".csv"))
  write.csv(results_NoPlat_TCGA_MRI[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_NoPlat_TCGA_MRI_Surv_", i, ".csv"))

  write.csv(results_Plat_NoRest[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_Plat_NoRest_LNM_", i, ".csv"))
  write.csv(results_Plat_NoRest[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_Plat_NoRest_Surv_", i, ".csv"))

  write.csv(results_Plat_TCGA_MRI[[i]]$LNM, paste0("4. Model_analysis/Transfer_results/Results_Plat_TCGA_MRI_LNM_", i, ".csv"))
  write.csv(results_Plat_TCGA_MRI[[i]]$Surv, paste0("4. Model_analysis/Transfer_results/Results_Plat_TCGA_MRI_Surv_", i, ".csv"))

  write.csv(testsets[[i]], paste0("4. Model_analysis/Transfer_results/Testset_", i, ".csv"))
}
```

```{r}

```

```{r}

```